CF Registry Analysis

### 2023-12-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry first draft survival done

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
execute:
  echo: false
  warning: false
editor: source
---

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(icenReg)
library(knitr)
library(eulerr)
library(broom)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
```

@@

```{r}
# Whole cohort
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
  data = s,cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018),break.x.by=1
)
kable(tidy(km_fit))
# By age group
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
  data = s,cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018),break.x.by=1
)
km_table <- tidy(km_fit)
km_table <- split.data.frame(km_table, km_table$strata)
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  kable(t, caption = name)
}))
```

@@

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
  "A_pulmonarycomplications1", "A_corticosteroids1", "liver_disease"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
ggsurvplot(survfit(cox_res),
  data = s, risk.table = T, cumevents = T, fun = "event",
  xlim = c(2003, 2018), break.x.by = 1
)
# Forest plot and coefficients table
ggforest(cox_res,data = s)
kable(tidy(cox_res))
# Diagnostics
test.ph= cox.zph(cox_res)
ggcoxzph(test.ph)
ggcoxdiagnostics(cox_res,type = "martingale")
```

>>>>

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    fig-width: 10
    fig-height: 10
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
execute:
  echo: false
  warning: false
editor: source
---

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(icenReg)
library(knitr)
library(eulerr)
library(broom)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
```

@@

```{r}
# Whole cohort
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
  data = s, cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
kable(tidy(km_fit))
# By age group
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
  data = s, cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
km_table <- tidy(km_fit)
km_table <- split.data.frame(km_table, km_table$strata)
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  kable(t, caption = name)
}))
```

@@

### Base model

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "tube_feeding", "liver_disease",
  "A_FEV1_pct_predicted", "NumPulmExacerbation",
  "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
ggsurvplot(survfit(cox_res),
  data = s, risk.table = T, cumevents = T, fun = "event",
  xlim = c(2003, 2018), break.x.by = 1
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
# test.ph= cox.zph(cox_res)
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

#### P. aeruginosa

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### Aspergillus

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### ABPA

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

### Vx770

```{r}
f <- update(f, . ~ . + A_Vx770)
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
ggsurvplot(survfit(cox_res),
  data = s, risk.table = T, cumevents = T, fun = "event",
  xlim = c(2003, 2018), break.x.by = 1
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
# test.ph= cox.zph(cox_res)
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

#### P. aeruginosa

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### Aspergillus

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### ABPA

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### Mycobacterium

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

### Vx809

```{r}
f <- update(f, . ~ . + A_VX809comb)
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
ggsurvplot(survfit(cox_res),
  data = s, risk.table = T, cumevents = T, fun = "event",
  xlim = c(2003, 2018), break.x.by = 1
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
# test.ph= cox.zph(cox_res)
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

#### P. aeruginosa

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### Aspergillus

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### ABPA

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```

#### Mycobacterium

```{r}
# Update formula
cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
  data = s, cluster = eDWID
)
# Forest plot and coefficients table
ggforest(cox_res, data = s)
```
